R tools for accessing research literature for text mining

### 2019-04-12 manuscript.tex
work on ms

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rOpenSci tools for accessing research literature for text mining},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{rOpenSci tools for accessing research literature for text mining}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{myrmecocystus(at)gmail.com} 
   \cortext[c1]{Corresponding author}
      \address[cstar]{rOpenSci, Museum of Paleontology, University of California, Berkeley,
CA, USA}

  \begin{abstract}
  Corresponding Author:

  Scott Chamberlain

  rOpenSci, Museum of Paleontology, University of California, Berkeley,
  CA, USA

  Email address:
  \href{mailto:myrmecocystus@gmail.com}{\nolinkurl{myrmecocystus@gmail.com}}

  Background. xxxx.

  Methods. xxxx.

  Results. xxxx.

  Discussion. xxxx.
  \end{abstract}

@@

There's more than 100 million articles published (source: Crossref API),
representing an enormous amount of knowledge. In addition to simply
reading these articles, they contain a vast trove of information of
interest to researchers for machine aided questions.

@@

Text-mining is the broad term associated with pulling information out of
articles. Given the importance of text-mining, good text-mining tools
are needed to make it easier for researchers. In particular, the R
programming language is used widely throughout many academic fields and
thus tools in R for text mining are of particular importance.

Here, we present an overview of text-mining tools in the R programming
language. We do not cover analysis tools per se, but rather those tools
for searching for, acquiring, and ``mashing up'' text.

@@

Those articles that are digital can be split into two groups: easily
machine readable and non-machine readable.

The machine readable articles are those in XML, JSON, or plain text
format. The former two, XML and JSON, are ideal for the machine readable
types because they are structured data, whereas plain text has no
structure - it's simply a set of characters with line breaks and spaces
in between.

Of the non-machine readable kind, there's PDFs. These can be broken out
into two groups: text based PDFs and scanned PDFs. The former are
converted from digital versions of various kinds (MS Word, OpenOffice,
markdown, etc.), while the latter are PDFs created by scanning in print
articles for which there is no digital version.

@@

A promising sign is an increasing number of open access publishers, yet
these represent a very small portion of the total articles (XXXXX)
(ref.).

With respect to paywalled articles, access varies by institution,
depending on what each institution decided to pay for. In addition, some
users may not realize access varies with IP address so that access from
campus vs.~from home (if not on a VPN) will drastically differ.

We can not end this section without mentioning SciHub. This is a last
resort option for many probably, providing access to full text of
articles that are normally paywalled. No tools in this manuscript
provide access to SciHub.

@@

xxx

\hypertarget{section}{%
\section{}\label{section}}

xxx

@@

xxxxx

@@

xxxxx

@@

xxxxx

@@

xxx

@@

The following is a breakdown of the major distinct functional parts of
\texttt{fulltext}.

@@

\texttt{ft\_search()} provides search access to nine different data
sources (PLOS, BMC, Crossref, Entrez, arXiv, bioRxiv, Europe PMC,
Scopus, Microsoft Academic), creating a mostly unified interface to all
data sources. The parts of each data source that are common are mostly
factored into the parameters of the \texttt{ft\_search()} function, and
we also allow the user to pass on data source specific options as
needed.

xxxxx

@@

\texttt{ft\_links()} provides two pathways to get links (URLs) for
articles, with a choice of four different data sources (PLOS, BMC,
Crossref, Entrez). First, you can use \texttt{ft\_search()}, then pass
the output of that function to \texttt{ft\_links()}. Second, you can
pass DOIs directly to \texttt{ft\_links()}. Both end up at the same
point, links for each article, if they could be found for the user
selected data source.

@@

\texttt{ft\_get()} fetchs full text of articles from many different data
sources. From the DOIs that are passed in to the function, we detect the
publisher, and there are specific plugins for certain publishers:

\begin{itemize}
\tightlist
\item
  aaas
\item
  aip
\item
  amersocclinoncol
\item
  amersocmicrobiol
\item
  arxiv
\item
  biorxiv
\item
  bmc
\item
  copernicus
\item
  crossref
\item
  elife
\item
  elsevier
\item
  entrez
\item
  frontiersin
\item
  ieee
\item
  informa
\item
  instinvestfil
\item
  jama
\item
  microbiology
\item
  peerj
\item
  pensoft
\item
  plos
\item
  pnas
\item
  royalsocchem
\item
  sciencedirect
\item
  scientificsocieties
\item
  wiley
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(corp)}
\CommentTok{#> Corpus consisting of 10 documents:}
\CommentTok{#> }
\CommentTok{#>    Text Types Tokens Sentences                    V1}
\CommentTok{#>   text1   143    262        10   10.2458/v22i1.21112}
\CommentTok{#>   text2   117    244         6   10.2458/v17i1.21696}
\CommentTok{#>   text3    75    118         4   10.2458/v25i1.23119}
\CommentTok{#>   text4     5      8         1    10.2458/v1i1.21154}
\CommentTok{#>   text5   105    171         7   10.1155/2011/868426}
\CommentTok{#>   text6   112    181         6   10.1155/2012/273413}
\CommentTok{#>   text7   117    240         8 10.5194/we-13-91-2013}
\CommentTok{#>   text8   140    245         9 10.5194/we-13-95-2013}
\CommentTok{#>   text9   107    202         7   10.1155/2014/198707}
\CommentTok{#>  text10   118    224         6   10.5402/2011/897578}
\CommentTok{#> }
\CommentTok{#> Source: /Users/sckott/github/ropensci/textmine/use-cases/* on x86_64 by sckott}
\CommentTok{#> Created: Fri Apr  5 11:36:04 2019}
\CommentTok{#> Notes:}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{(res1 <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \StringTok{'plos'}\NormalTok{))}
\CommentTok{#> Query:}
\CommentTok{#>   [ecology] }
\CommentTok{#> Found:}
\CommentTok{#>   [PLoS: 47337; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0] }
\CommentTok{#> Returned:}
\CommentTok{#>   [PLoS: 10; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0]}
\end{Highlighting}
\end{Shaded}

@@

@@

\hypertarget{references}{%
\section{References}\label{references}}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={R tools for accessing research literature for text mining},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{R tools for accessing research literature for text mining}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{myrmecocystus(at)gmail.com} 
   \cortext[c1]{Corresponding author}
      \address[cstar]{rOpenSci, Museum of Paleontology, University of California, Berkeley,
CA, USA}

  \begin{abstract}
  Text-mining is a powerful method for answering research questions.
  However, getting texts to extract information can be a daunting and
  complicated task. The primary reason for this is publishers. There are
  thousands of different publishers, each with their own licenses, URL
  patterns, access options, and more. Layered on top of that is the varied
  access each user has based on their institutional affiliation. Here, I
  introduce a suite of software packages in the R programming language for
  fetching texts. The tapestry of different publishers, access levels, and
  other factors requires a patchwork of approaches for getting texts to
  users. The flagship R package called fulltext attempts to simplify
  search and retrieval of texts for text-mining. The fulltext package,
  along with many others, make acquiring texts easier than ever,
  facilitating answering research questions with text-mining.
  \end{abstract}

@@

There's more than 100 million research articles published (Crossref API:
\url{https://github.com/CrossRef/rest-api-doc}), representing an
enormous amount of knowledge. In addition to simply reading these
articles, the articles contain a vast trove of information of interest
to researchers for machine aided questions (Kong \& Gerstein, 2018; Usai
et al., 2018).

@@

Text-mining is the broad term associated with pulling information out of
articles. Given the importance of text-mining, good text-mining tools
are needed to make it easier for researchers to do. Graphical user
interface (GUI) based text mining tools are available (e.g., Ba \&
Bossy, 2016) and some research papers have used them (Chaix et al.,
2019), but given the urgent recent call to action for more reproducible
research (Open Science Collaboration, 2015; Camerer et al., 2016, 2018),
we must move away from GUI based tools as fast as possible.

In particular, the R programming language is used widely throughout many
academic fields and thus tools in R for text mining are of particular
importance because they can be adopted by academics rapidly.

Here, I present an overview of text-mining tools in the R programming
language, not for text-mining analysis, but rather those tools for
searching for, acquiring, and extracting particular chunks of texts
(e.g., title, abstract, authors). Most of the packages are part of the
rOpenSci suite (\url{https://ropensci.org/}).

@@

Those articles that are digital (which in theory includes all articles)
can be split into two groups: machine readable and non-machine readable.

The machine readable articles are those in XML\footnote{\url{https://www.w3.org/TR/xml/}},
JSON\footnote{\url{https://tools.ietf.org/html/rfc7159}}, or plain text
format. The former two, XML and JSON, are the best machine readable
types because they are structured data\footnote{\url{https://en.wikipedia.org/wiki/Data_model}},
whereas plain text has no structure - it's simply a set of characters
with line breaks and spaces in between.

Of the non-machine readable types, the most noteable is the Portable
Document Format (PDF)\footnote{\url{https://en.wikipedia.org/wiki/PDF}}.
These can be broken out into two groups: text based PDFs and scanned
PDFs. The former are converted from digital versions of various kinds
(MS Word, OpenOffice, LaTeX, markdown, etc.), while the latter are
created by scanning print articles to a PDF format. Text-based PDFs are
much better for text-mining purposes as plain text can be extracted
easily in R with \href{https://github.com/ropensci/pdftools}{pdftools},
a binding to \href{https://poppler.freedesktop.org/}{libpoppler}.
However, with scanned PDFs, text must be extracted using Optimal
Character Recognition (OCR; see R package
\href{https://github.com/ropensci/tesseract}{tesseract}), which isn't
always a clean solution, especially compared to true text based PDFs.

The reality in scholarly publishing is all publishers, if they provide
any access to their articles, only provide PDF format. Very few
publishers, with some quite large (Elsevier, Pensoft, PLOS), provide XML
format. Although most publishers most likely have the XML behind each of
their articles, they for some indefensible reason do not share it -
making text-mining more difficult. Some provide plain text (Elsevier). I
only know of one publisher that provides full text as JSON (PLOS). Thus,
text-mining, in most cases, will require extracting text from PDFs.

@@

A promising sign is an increasing number of open access articles, yet
open access articles represent a small percent of all articles: an
estimate in 2018 said that 28\% of the scholarly literature was open
access (Piwowar et al., 2018).

With respect to paywalled articles, access varies by institution,
depending on each institution's publisher contracts. MORE ABOUT THIS
\ldots{}

Some may not realize access to articles varies with IP address so that
access from campus vs.~from home (if not on a VPN) will drastically
differ. Sometimes a VPN is required, and this can provide a significant
technical hurdle to users attempting to do text-mining work.

One final hurdle in text-mining comes unsurprisingly from Elsevier. They
use so-called ``fences'' for programmatic access. That is, even if a
person trying to get an article programmatically their institution has
access to and they have access to, and they are on the correct IP
address, they may still not get access to an Elsevier article. Elsevier
puts in place these fences and only if you contact their technical team
directly can you get these fences removed, and only then on a per
institution basis.

I can not end this section without mentioning SciHub. This is a last
resort option for many probably (or possibly first, depending on your
level of access), providing access to full text of articles that are
normally paywalled. No tools in this manuscript provide access to
SciHub.

@@

A text-mining project starts with a question. From that question,
researchers then attempt to acquire scholarly articles for text-mining.
Finding those articles is not altogether straight-forward.

There are many places to search for articles; a non-exhaustive list:
Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect, Web
of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access
Journals, Open Knowledge Maps, and more. It's probably difficult to know
where the best place is to search. Some of these are paywalled (e.g.,
Web of Science), and some are not.

The most important aspect about any source for article search with
respect to reproducible research is being able to use the data source
programmatically. Of those listed above, the following can be used
programmatically: Microsoft Academic Research, Scopus, ScienceDirect,
Pubmed/Entrez, Europe PMC, and Directory of Open Access Journals. All of
these are included in the R package
\href{https://github.com/ropensci/fulltext}{fulltext}, discussed further
below.

@@

Directory of Open Access Journals (DOAJ) maintains data on open access
journals, as well as some portion of the articles in those journals.
Thus, you can search for journals as well as articles with DOAJ. The R
interface for DOAJ is \href{https://github.com/ropensci/jaod}{jaod}.

@@

JSTOR's Data for Research program gives institutions with access to
JSTOR, access to full text of articles within JSTOR. There is no way
however to make the interaction with JSTOR completely programmatic, thus
making reproducible research very difficult. Nonetheless, there is an R
package (\href{https://github.com/ropensci/jstor}{jstor}) for using data
from JSTOR's Data for Research.

@@

ORCID (\url{https://orcid.org/}) is an organization keeping track of
identifiers and metadata for researchers around the world. Individuals
can optionally maintain metadata on their scholarly works connected to
their account with ORCID. Thus, across all of ORCID, a significant cache
of metadata is accruing on scholarly works, their funding amounts,
collaborators, etc., useful for bibliometrics research and more. The R
interface for ORCID is
\href{https://github.com/ropensci/rorcid}{rorcid}.

@@

The Open Citation Corpus (\url{http://opencitations.net/}) holds records
of which articles cite which other articles, allowing for all important
research on the scholarly web of citation. Citation data has been very
closely guarded until recently, but the largest publishers are still not
contributing to the Open Citation Corpus. The R interface to the Open
Citation Corpus is
\href{https://github.com/ropenscilabs/rcitoid}{rcitoid}.

@@

The following is a breakdown of the major distinct parts of
\texttt{fulltext}.

@@

\texttt{ft\_search()} provides search access to nine different data
sources (PLOS, BMC, Crossref, Entrez, arXiv, bioRxiv, Europe PMC,
Scopus, Microsoft Academic), creating a mostly unified interface to all
data sources. The parts of each data source that are common are for the
most part factored out into the parameters of the \texttt{ft\_search()}
function: query term(s), pagination (number of results, result number to
start at). In addition, we allow the user to pass on data source
specific options to refine the search per data source.

With \texttt{ft\_search()}, you can query any combination of the nine
data sources at once. The returned object is a list, with access to
results of each data source by its name (e.g., \texttt{\$plos}, or
\texttt{\$crossref}). For each data source, the returned object does
vary because the returned data from each data source widely varies; for
the most part data.frame's are returned. For those data sources not
queried, their slot is empty.

One important aspect of the research result we highlight is the licenses
in the returned data for each data source.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \KeywordTok{c}\NormalTok{(}\StringTok{"plos"}\NormalTok{, }\StringTok{"crossref"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

The results for this PLOS search have all CC-BY licenses

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x}\OperatorTok{$}\NormalTok{plos}
\CommentTok{#> Query: [ecology] }
\CommentTok{#> Records found, returned: [47257, 10] }
\CommentTok{#> License: [CC-BY] }
\CommentTok{#>                              id}
\CommentTok{#> 1  10.1371/journal.pone.0001248}
\CommentTok{#> 2  10.1371/journal.pone.0059813}
\CommentTok{#> 3  10.1371/journal.pone.0080763}
\CommentTok{#> 4  10.1371/journal.pone.0155019}
\CommentTok{#> 5  10.1371/journal.pone.0175014}
\CommentTok{#> 6  10.1371/journal.pone.0150648}
\CommentTok{#> 7  10.1371/journal.pone.0208370}
\CommentTok{#> 8  10.1371/journal.pcbi.1003594}
\CommentTok{#> 9  10.1371/journal.pone.0102437}
\CommentTok{#> 10 10.1371/journal.pone.0166559}
\end{Highlighting}
\end{Shaded}

Whereas the results for this Crossref search have mixed licenses

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x}\OperatorTok{$}\NormalTok{crossref}
\CommentTok{#> Query: [ecology] }
\CommentTok{#> Records found, returned: [164657, 10] }
\CommentTok{#> License: [variable, see individual records] }
\CommentTok{#>    archive               container.title    created  deposited}
\CommentTok{#> 1  Portico                       Ecology 2006-05-03 2018-08-04}
\CommentTok{#> 2  Portico                       Ecology 2006-05-03 2018-08-04}
\CommentTok{#> 3       NA                       Ecology 2006-05-03 2018-08-04}
\CommentTok{#> 4       NA                       Ecology 2006-05-03 2018-08-04}
\CommentTok{#> 5       NA                       Ecology 2006-05-03 2018-08-04}
\CommentTok{#> 6       NA                       Ecology 2006-05-03 2018-08-04}
\CommentTok{#> 7       NA                       Ecology 2006-05-09 2018-08-01}
\CommentTok{#> 8  Portico                       Ecology 2017-04-26 2019-03-08}
\CommentTok{#> 9       NA Trends in Ecology & Evolution 2002-07-25 2017-06-14}
\CommentTok{#> 10      NA Journal of Industrial Ecology 2014-11-21 2017-06-23}
\CommentTok{#> Variables not shown: published.print (chr), published.online (chr), doi}
\CommentTok{#>      (chr), indexed (chr), issn (chr), issue (chr), issued (chr), member}
\CommentTok{#>      (chr), page (chr), prefix (chr), publisher (chr), reference.count}
\CommentTok{#>      (chr), score (chr), source (chr), title (chr), type (chr), url (chr),}
\CommentTok{#>      volume (chr), author (list), link (list), license (list), subject}
\CommentTok{#>      (chr), alternative.id (chr), subtitle (chr), reference (list)}
\end{Highlighting}
\end{Shaded}

You can dig into the license field for each article, with URLs holding
information on each license

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{vapply}\NormalTok{(x}\OperatorTok{$}\NormalTok{crossref}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{license, }\ControlFlowTok{function}\NormalTok{(w) w}\OperatorTok{$}\NormalTok{URL[}\DecValTok{1}\NormalTok{], }\StringTok{""}\NormalTok{)}
\CommentTok{#>  [1] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
\CommentTok{#>  [2] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
\CommentTok{#>  [3] "http://doi.wiley.com/10.1002/tdm_license_1"  }
\CommentTok{#>  [4] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
\CommentTok{#>  [5] "http://doi.wiley.com/10.1002/tdm_license_1"  }
\CommentTok{#>  [6] "http://doi.wiley.com/10.1002/tdm_license_1"  }
\CommentTok{#>  [7] "http://doi.wiley.com/10.1002/tdm_license_1"  }
\CommentTok{#>  [8] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
\CommentTok{#>  [9] "http://www.elsevier.com/tdm/userlicense/1.0/"}
\CommentTok{#> [10] "http://doi.wiley.com/10.1002/tdm_license_1.1"}
\end{Highlighting}
\end{Shaded}

@@

\texttt{ft\_links()} provides two pathways to get links (URLs) for
articles, with a choice of four different data sources (PLOS, BMC,
Crossref, Entrez). First, you can use \texttt{ft\_search()}, then pass
the output of that function to \texttt{ft\_links()}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{"ecology"}\NormalTok{, }\DataTypeTok{from =} \StringTok{"entrez"}\NormalTok{)}
\KeywordTok{ft_links}\NormalTok{(out)}
\CommentTok{#> <fulltext links>}
\CommentTok{#> [Found] 6 }
\CommentTok{#> [IDs] ID_30964001 ID_30962485 ID_30962432 ID_30952928 ID_30674747}
\CommentTok{#>      ID_30674743 ...}
\end{Highlighting}
\end{Shaded}

Second, you can pass DOIs directly to \texttt{ft\_links()}. Both end up
at the same point, links for each article, if they could be found for
the user selected data source.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# }\AlertTok{FIXME}
\KeywordTok{ft_links}\NormalTok{(out}\OperatorTok{$}\NormalTok{entrez}\OperatorTok{$}\NormalTok{data}\OperatorTok{$}\NormalTok{doi)}
\end{Highlighting}
\end{Shaded}

@@

\texttt{ft\_get()} fetchs full text of articles from many different data
sources. From the DOIs that are passed in to the function, we detect the
publisher, and there are specific plugins for certain publishers: AAAS,
American Institute of Physics, American Society of Clinical Oncology,
American Society for Microbiology, arXiv, bioRxiv, BiomedCentral,
Copernicus, Crossref, Elife, Elsevier, Pubmed/PMC via NCBI's Entrez,
Frontiers, IEEE, Informa, Instituto de Investigaciones Filologicas,
American Medical Association, Microbiology Society, PeerJ, Pensoft,
PLOS, PNAS, Royal Society of Chemistry, ScienceDirect, Scientific
Societies, and Wiley.

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(corp)}
\CommentTok{#> Corpus consisting of 10 documents:}
\CommentTok{#> }
\CommentTok{#>    Text Types Tokens Sentences                    V1}
\CommentTok{#>   text1   143    262        10   10.2458/v22i1.21112}
\CommentTok{#>   text2   117    244         6   10.2458/v17i1.21696}
\CommentTok{#>   text3    75    118         4   10.2458/v25i1.23119}
\CommentTok{#>   text4     5      8         1    10.2458/v1i1.21154}
\CommentTok{#>   text5   105    171         7   10.1155/2011/868426}
\CommentTok{#>   text6   112    181         6   10.1155/2012/273413}
\CommentTok{#>   text7   117    240         8 10.5194/we-13-91-2013}
\CommentTok{#>   text8   140    245         9 10.5194/we-13-95-2013}
\CommentTok{#>   text9   107    202         7   10.1155/2014/198707}
\CommentTok{#>  text10   118    224         6   10.5402/2011/897578}
\CommentTok{#> }
\CommentTok{#> Source: /Users/sckott/github/ropensci/textmine/use-cases/* on x86_64 by sckott}
\CommentTok{#> Created: Thu Apr 11 11:20:19 2019}
\CommentTok{#> Notes:}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{(res1 <-}\StringTok{ }\KeywordTok{ft_search}\NormalTok{(}\DataTypeTok{query =} \StringTok{'ecology'}\NormalTok{, }\DataTypeTok{from =} \StringTok{'plos'}\NormalTok{))}
\CommentTok{#> Query:}
\CommentTok{#>   [ecology] }
\CommentTok{#> Found:}
\CommentTok{#>   [PLoS: 47272; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0] }
\CommentTok{#> Returned:}
\CommentTok{#>   [PLoS: 10; BMC: 0; Crossref: 0; Entrez: 0; arxiv: 0; biorxiv: 0; Europe PMC: 0; Scopus: 0; Microsoft: 0]}
\end{Highlighting}
\end{Shaded}

@@

To make text mining easier:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  publishers should provide XML if they have it
\item
  publishers should not change URL patterns so often, or at all
\item
  publishers should keep their Crossref metadata up to date
\item
  remove all paywalls (Easy, yes?)
\item
  \ldots{}
\end{enumerate}

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}
